Learning Auditory Similarities Between Instrumental Playing Techniques

### 2019-10-28 paper/paper.tex

Music information retrieval operates at either of two levels: symbolic and auditory \cite{downie2003mir}.
By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, or intensity at the source.
Timbre, in contrast, is a qualitative attribute of music, thus irreducible to a one-dimensional axis \cite{siedenburg2019chapter}.
As a result, symbolic representations describe timbre indirectly: either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked \cite{kostka2016book}).

Despite their widespread use, purely linguistic references to timbre fail to convey the intention of the composer.
On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them \cite{antoine2018isma}.
On the other hand, the sole mention of a playing technique does not determine any intended effect; and therefore, fails to generalize to the entire instrumentarium \cite{kolozali2011ismir}.
For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.

The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain.
Instead, it involves a cognitive process which arises from the subjective experience of listening \cite{erickson1975book}.
The task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
Rather than to merely discriminate instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters --- \ie{} duration, pitch, and intensity --- being equal \cite{thoret2018jasa}.

@@

This article proposes a machine listening system which computes the dissimilarity in timbre between any two audio samples.
Our system composes two stages: unsupervised feature extraction and supervised metric learning.
The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN) \cite{weinberger2009distance}.
It recombines scattering coefficients so as to fit a training set of human judgments \cite{mcadams1995psychres}, which may either stem from the perceptual idiosyncracies of a single subject or from a collective consensus.

This article gathers three novel contributions.
First, our model encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production \cite{lostanlen2018extended}.
Specifically, we fit pairwise judgments for 78 different techniques arising from 16 instruments, some of which include removable timbre-altering devices such as mutes.
Secondly, we purposefully disregard the playing technique metadata underlying each audio sample during the training phase of our model.
Said otherwise, we rely on listeners, not performers, to define and evaluate the task at hand.
A comprehensive ablation study of our system demonstrates that all design choices presented hereafter benefit the retrieval precision of our system.
Thirdly, we supplement our quantitative benchmark with some visualizations of time--frequency scattering coefficients in the rate--scale domain, for various typical samples of realistic instrumental playing techniques.
These visualizations confirm the theoretical predictions of previous research \cite{anden2015mlsp}, which relied on simple synthetic sounds such as exponential chirps.
They are also in line with visualizations of the modulation power spectrum in auditory neurophysiology \cite{patil2012ploscompbiol}, while offering an accelerated algorithm for scalable feature extraction.

Section \ref{sec:soa} reviews recent research in the computational modeling of musical timbre.
Section \ref{sec:survey} describes our data collection procedure for auditory similarities across playing techniques.
Section \ref{sec:method} presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
Section \ref{sec:results} reports the results of an ablation study of our system.
Section \ref{sec:discussion} discusses the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects: \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp

@@

On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.
Because such tags often belong to the metadata of music releases, the process of curating a training set for musical instrument classification requires little or no human intervention.
Although scraping user-generated content from online music hosting platforms may not always reflect the true instrumentation with perfect accuracy, it offers a scalable and ecologically valid insight onto the acoustic underpinnings of musical timbre.
Furthermore, supplementing user-generated content with the outcome of a crowdsourced annotation campaign allows an explicit verification of instrument tags.
For instance, the Open-MIC dataset \cite{humphrey2018ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of 20k polyphonic music recordings spanning across 20 instruments, as a derivative of the Free Music Archive (FMA) dataset \cite{defferrard2017ismir}.
Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from 8 instruments, as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
Over the past decade, the availability of large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning architectures, and notably deep convolutional networks, in music instrument recognition \cite{mcfee2015ismir,pons2017eusipco,gururani2018ismir}.
Notwithstanding the growing accuracy of these architectures in the large data regime, it remains unclear how to extend them from musical instrument recognition to playing technique recognition, where labeled samples are considerably scarcer.
We refer to \cite{han2017taslp} for a recent review of the state of the art in this domain.

On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.
Rather, each excerpt must be played back to multiple human listeners.
Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, thus restricting the size of the musical corpus under study.
Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in artificial neural networks.
Nevertheless, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}.
The STRF of a neuron is a function of time and frequency which represents the optimal predictor of its post-stimulus time histogram (PSTH) during exposure to a diverse range of auditory stimuli \cite{aertsen1981biolcyb}.
The simplest method to compute it in practice is by reverse correlation, i.e. by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but eliciting cross-term interferences \cite{eggermont1993hearing}.
Since then, the STRF of a neuron was redefined as a spectrographic representation of its spike-triggered average \cite{klein2000compneur}.
Although this new definition is necessarily tied to a choice of spectrogram parameters, it yields more interpretable patterns than a Wigner-Ville distribution.
In particular, a substantial portion of spectrographic STRFs exhibits a ripple-like response around a given region $(t, \lambda)$ of the time--frequency domain.
This response can be approximately described by a pair of scalar values: a temporal modulation rate $\alpha$ in Hertz and a frequential modulation rate $\beta$ in cycles per octave.
Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the selected center time $t$ and center frequency $\lambda$.
This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model'' \cite{chi2005jasa}, which densely covers the rate--scale domain $(\alpha, \beta)$.

@@

The reasons above make STRFs an appealing feature extractor for a perceptual description of timbral similarity across instrumental playing techniques.
Nonetheless, current implementations of STRF suffer from a lack of scalability, which explains why they have found few applications in MIR thus far.
Indeed, the full cortical model is usually computed via two-dimensional Fourier transforms over adjacent time--frequency regions, followed by averaging around specific rates and scales.
On the contrary, joint time--frequency scattering offers a faster extraction of spectrotemporal modulations while preserving properties of differentiability \cite{andreux2019jmlr} and invertibility \cite{lostanlen2019dafx}.
Such acceleration is made possible by discretizing the wavelet transforms involved in time--frequency scattering according to a multirate scheme, both along the time and the log-frequency variables \cite{anden2019tsp}.
This multirate scheme also has the benefit of drawing an explicit connection between scattering networks and deep convolutional networks, as both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class \cite{anden2019tsp}.
Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features, and achieved state-of-the-art results on a dataset of continuous performance recordings of the Chinese bamboo flute, involving periodic spectrotemporal modulations such as vibratos and trills.
Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be known whether this finding can scale up from a single instrument to an entire orchestra.
In addition, all previous publications in time--frequency scattering, whether separable or joint, lack a human-centric evaluation, remotely from any classification task.

The contributions of this paper strive to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
From the standpoint of MIR, the model presented here offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain, whether random or adversarial.
Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate \emph{in silico} for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.
% enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
%We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.

@@

The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
In this article, we focus on the former three, and leave the question of learning auditory similarities between percussion instruments to future research.
The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey) these are rarely shared with pitched instruments.
In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
We refer to \cite{x} and \cite{x} for a review of recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively. % TODO VL CITE

@@

% one paragraph on annotation interface
Thirdly,
% in 30 to 60 minutes.

@@

\ml{introduce the system globally, while stressing the ability of the system to be conditioned for one or many people (VL). consensus ou un compositeur particulier.}

The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
Furthermore, we may calculate the scattering tranforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014deep}.
They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.

The standard Euclidean distance on the raw scattering coefficients  provides some sort universal representation of timbre dissimilarity. Considering some sort of learning to apply further transformations to the scattering coefficients and compute the Euclidean distance on the result.

@@

The processing pipeline presented in Figure \ref{fig:pipeline} addresses the problem of conditioning a timbre space using perceptual judgments of timbral similarity. The relevance of the many design choices that had to be made  are addressed in this section using an experimental protocol that considers a dataset of audio that will be given as input to the processing pipeline and one metric used for evaluating its performance under a given experimental setting.

@@

The perceptual data gathered during the perceptual experiment considers 78 \ipts each represented by one sound recording that the subject could listen to before clustering the \ipts. For those 78 \ipts, 31 clusterings $C_k, 0<k<31, Card(C_k)=78$ are available, that are our reference in terms of timbre similarity. If \ipt $i$ and \ipt $j$ are clustered together by subject $k$, then $C_k(i)=C_k(j)$. In essence, an effective ranking system should retrieve sounds that belong to \ipts that are clustered together in each of the 31 clusterings. Depending on the preferences of subjects, there may be competition between clusterings that should be mitigated when appreciating quantitatively the ranking performance.

In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects. Assuming invariance to timbre with respect to pitch and nuance is in line with the ansi definition of timbre \ml{vl, citation. reformuler tout le paragraphe}, it is  thus taken in this paper as a reasonable assumption. Quantification of this invariance is left for further study.

The whole range of nuance and pitch is thus considered for each of the 78 \ipts, leading to a dataset composed of 9346 sound recordings. The clusterings $C_k$ are propagated to this dataset to produce new reference clusterings $D_k, 0<k<31, Card(D_k)=9346$ using the following scheme. If a sound of this dataset $s_a$ corresponds to \ipt $i$, then $D_k(a) = C_k(i), \forall k$.

@@

The metric considered in this paper is the average precision at rank $k=5$ (ap@5). It allows us to evaluate if the outputs of the proposed system are likely to satisfy the user. The ap@5 is computed as follows.

@@

For a given reference clustering $D_k$, the p@5 is computed as follows. For each sound $s_q, 0<q<9346$ considered as a query, its 5 nearest neighbors $s_n, 0<n<5$ in the timbre space under evaluation are retrieved using the Euclidean distance between corresponding features vectors. For this query, the precision is defined as:
$$
p@5_q = \frac{\sum_{n=5}^5 D_q==D_n}{5}
$$
The p@5 is the average $p@5_q$ computed over all queries.

@@

The choice of the rank is chosen with the application in mind. When searching for sounds that are similar to a query, the listener is likely to listen to up to $5$ proposals, rarely more. For the same reason, more global metrics commonly considered for ranking tasks such as the mean average precision is not considered here.

@@

The processing pipeline is composed of 2 main building blocks. The first block $B_1$ takes a digital audio recording as input and outputs one vector of features. For this block, the main factors are the time support $B_1^t$ for computing the features and the type of features $B_1^f$. We considered $B_1^t \in \{ 25, 128, 250, 500, 1000 \}$ ms and $B_1^f \in \{ \text{mfccs}, \text{separable scattering}, \text{joint scattering} \}$. The mfccs are the standard Mel frequency cepstral coefficients with 40 Mel bands and selection of the 13$^\text{th}$ first discrete cosine transform (dct) coefficients.

@@

As found in many studies, log compression and median processing are beneficial post processing steps for the scattering.

For each sound, a set of features are computed for each time support. This set is averaged over time to give one feature vector per sound.
All features are standardized with zero mean and unit variance across the dataset.

>>>>

Music information retrieval (MIR) operates at either of two levels: symbolic and auditory \cite{downie2003mir}.
By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, and intensity at the source.
Timbre, in contrast, is a qualitative attribute of music, and is thus not reducible to a one-dimensional axis \cite{siedenburg2019chapter}.
As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked \cite{kostka2016book}).

Despite their widespread use, purely linguistic references to timbre fail to convey the intention of the composer.
On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them \cite{antoine2018isma}.
On the other hand, the sole mention of a playing technique does not determine any intended effect; as a result, it fails to generalize to the entire instrumentarium \cite{kolozali2011ismir}.
For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.

The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain.
Instead, it involves a cognitive process which arises from the subjective experience of listening \cite{erickson1975book}.
The task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
Rather than to merely discriminate instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters---duration, pitch, and intensity---being equal \cite{thoret2018jasa}.

@@

This article proposes a machine listening system which computes the dissimilarity in timbre between any two audio samples.
Our system consists of two stages: unsupervised feature extraction and supervised metric learning.
The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp,anden2019tsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN) \cite{weinberger2009distance}.
It recombines scattering coefficients so as to fit a training set of human judgments \cite{mcadams1995psychres}, which may either stem from the perceptual idiosyncracies of a single subject or from a collective consensus.

This article gathers three novel contributions.
First, our model encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production \cite{lostanlen2018extended}.
Specifically, we fit pairwise judgments for 78 different techniques arising from 16 instruments, some of which include removable timbre-altering devices such as mutes.
Secondly, we purposefully disregard the playing technique metadata underlying each audio sample during the training phase of our model.
In other words, we rely on listeners, not performers, to define and evaluate the task at hand.
A comprehensive ablation study of our system demonstrates that all design choices presented hereafter benefit the retrieval precision of our system.
Thirdly, we supplement our quantitative benchmark with some visualizations of time--frequency scattering coefficients in the rate--scale domain, for various typical samples of realistic instrumental playing techniques.
These visualizations confirm the theoretical predictions of previous research \cite{anden2019tsp}, which relied on simple synthetic sounds such as exponential chirps.
They are also in line with visualizations of the modulation power spectrum in auditory neurophysiology \cite{patil2012ploscompbiol}, while offering an accelerated algorithm for scalable feature extraction.

\ja{These section numbers aren't showing up since we're not numbering the sections (per the style guide).}
Section \ref{sec:soa} reviews recent research in the computational modeling of musical timbre.
Section \ref{sec:survey} describes our data collection procedure for auditory similarities across playing techniques.
Section \ref{sec:method} presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
Section \ref{sec:results} reports the results of an ablation study of our system.
Section \ref{sec:discussion} discusses the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects: \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp

@@

On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags,'' encoded as binary outputs within some predefined label space.
Because such tags often belong to the metadata of music releases, the process of curating a training set for musical instrument classification requires little or no human intervention.
Although scraping user-generated content from online music hosting platforms may not always reflect the true instrumentation with perfect accuracy, it offers a scalable and ecologically valid insight onto the acoustic underpinnings of musical timbre.
Furthermore, supplementing user-generated content with the outcome of a crowdsourced annotation campaign allows an explicit verification of instrument tags.
For instance, the Open-MIC dataset \cite{humphrey2018ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of 20000 polyphonic music recordings spanning 20 instruments as a derivative of the Free Music Archive (FMA) dataset \cite{defferrard2017ismir}.
Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from 8 instruments as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
Over the past decade, the availability of large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning architectures, and notably deep convolutional networks, in music instrument recognition \cite{mcfee2015ismir,pons2017eusipco,gururani2018ismir}.
Notwithstanding the growing accuracy of these architectures in the large data regime, it remains unclear how to extend them from musical instrument recognition to playing technique recognition, where labeled samples are considerably scarcer.
We refer to \cite{han2017taslp} for a recent review of the state of the art in this domain.

On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.
Rather, each excerpt must be played back to multiple human listeners.
Yet collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, restricting the size of the musical corpus under study.
Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in deep neural networks.
Nevertheless, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
\ja{This is quite an abrupt transition.
Why ``nevertheless''?}%
This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}.
The STRF of a neuron is a function of time and frequency which represents the optimal predictor of its post-stimulus time histogram (PSTH) during exposure to a diverse range of auditory stimuli \cite{aertsen1981biolcyb}.
The simplest method to compute it in practice is by reverse correlation, i.e.\ by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but eliciting cross-term interferences \cite{eggermont1993hearing}.
Since then, the STRF of a neuron was redefined as a spectrographic representation of its spike-triggered average \cite{klein2000compneur}.
Although this new definition is necessarily tied to a choice of spectrogram parameters, it yields more interpretable patterns than a Wigner--Ville distribution.
In particular, a substantial portion of spectrographic STRFs exhibits a ripple-like response around a given region $(t, \lambda)$ of the time--frequency domain.
\ja{``Spectrographic''?}
This response can be approximately described by a pair of scalar values: a temporal modulation rate $\alpha$ in Hertz and a frequential modulation rate $\beta$ in cycles per octave.
Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the selected center time $t$ and center frequency $\lambda$.
This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model'' \cite{chi2005jasa}, which densely covers the rate--scale domain $(\alpha, \beta)$.

@@

The reasons above make STRFs an appealing feature extractor for a perceptual description of timbral similarity across instrumental playing techniques.
Nonetheless, current implementations of STRF suffer from a lack of scalability, which explains why they have found few applications in MIR thus far.
Indeed, the full cortical model is usually computed via two-dimensional Fourier transforms over adjacent time--frequency regions, followed by averaging around specific rates and scales.
On the contrary, joint time--frequency scattering offers a faster extraction of spectrotemporal modulations while preserving properties of differentiability \cite{andreux2019jmlr} and invertibility \cite{lostanlen2019dafx}.
Such acceleration is made possible by discretizing the wavelet transforms involved in time--frequency scattering according to a multirate scheme, both along the time and the log-frequency variables \cite{anden2019tsp}.
This multirate scheme also has the benefit of drawing an explicit connection between scattering networks and deep convolutional networks, as both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class \cite{anden2019tsp}.
Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features \cite{anden2014deep}, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be seen whether these findings can scale up from a single instrument to an entire orchestra.
In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation, remotely from any classification task.
\ja{I don't understand this last sentence.
What does ``remotely'' mean here?}%

The contributions of this paper strive to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
From the standpoint of MIR, the model presented here offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain.
Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.

\ja{This section is way too long.
It's important that we situate ourselves with respect to the existing literature, but we don't need to spend 2+ pages doing so.
This currently reads more like a laundry list of various work related to timbre perception on musical instruments on one hand and a history of STRFs on the other.
We need to keep in mind what our central themes (wider range of timbre, similarity-based evaluation, simple scattering+linear model) are and link the existing work to them in a concise manner.
}

% enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
%We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.

Following the above discussed matters, we propose a ranking system that is based on a processing pipeline made of two main components, see Figure \ref{fig:pipeline}.
The first component maps the audio signal of the query and the items of the database into a high-dimensional timbre space using the scattering transform.
The second component warps this timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in Section \ref{sec:survey}.
This warping, or reweighting, of the space is performed using LMNN.
\ja{Again, section numbers are not showing up here.}%

The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
Furthermore, we may calculate the scattering transforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014deep}.
They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
\ja{Need to flesh out the above with some equations.
We're no longer talking to biologists, so we don't need to be afraid.}%

The standard Euclidean distance on the raw scattering coefficients provides a universal representation of timbre dissimilarity.
We now consider a transformation of this distance, in effect warping the timbre space, in order to more closely mimic a particular set of timbre similarity judgments.

@@

\section*{Timbre similarity judgments}
\label{sec:survey}
The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
In this article, we focus on the former three, and leave the question of learning auditory similarities between percussion instruments to future research.
The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey) these are rarely shared with pitched instruments.
In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
We refer to \cite{x} and \cite{x} for a review of recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively. % TODO VL CITE

@@

% one paragraph on annotation interface

Thirdly,
% in 30 to 60 minutes.

@@

@@

The processing pipeline presented in Figure \ref{fig:pipeline} addresses the problem of conditioning a timbre space using perceptual judgments of timbral similarity.
The relevance of the many design choices that had to be made are addressed in this section using an experimental protocol that considers a dataset of audio that will be given as input to the processing pipeline and one metric used for evaluating its performance under a given experimental setting.
\ja{This last sentence is quite vague.
What are we trying to say?}%

@@

The perceptual data gathered during the perceptual experiment considers 78 \ipts, each represented by one sound recording that the subject could listen to before clustering the \ipts.
For those 78 \ipts, 31 clusterings $C_k, 0<k<31, \Card(C_k)=78$ are available, that are our reference in terms of timbre similarity.
\ja{What is $C_k$ here?
Also, should it be $0 < k \le 31$?
Might be more natural to say that $k = 1, 2, \ldots, 31$, in that case.}%
If \ipt $i$ and \ipt $j$ are clustered together by subject $k$, then $C_k(i)=C_k(j)$.
In essence, an effective ranking system should retrieve sounds that belong to \ipts that are clustered together in each of the 31 clusterings.
Depending on the preferences of subjects, there may be competition between clusterings that should be mitigated when appreciating quantitatively the ranking performance.

In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects.
Assuming invariance to timbre with respect to pitch and nuance is in line with the ANSI definition of timbre \ml{vl, citation. reformuler tout le paragraphe}, it is thus taken in this work as a reasonable assumption.
Quantification of this invariance is left for further study.

The whole range of nuance and pitch is thus considered for each of the 78 \ipts, leading to a dataset composed of 9346 sound recordings.
The clusterings $C_k$ are propagated to this dataset to produce new reference clusterings $D_k, 0<k<31, \Card(D_k)=9346$ using the following scheme.
If a sound of this dataset $s_a$ corresponds to \ipt $i$, then $D_k(a) = C_k(i), \forall k$.
\ja{What's $s_a$ and $a$ here?
We need to be explicit about these definitions.}%

@@

The metric considered in this work is the average precision at rank $k=5$ (ap@5).
It allows us to evaluate if the outputs of the proposed system are likely to satisfy the user.
The ap@5 is computed as follows.

@@

For a given reference clustering $D_k$, the p@5 is computed as follows.
For each sound $s_q, 0<q<9346$ considered as a query, its 5 nearest neighbors $s_n, 0<n<5$ in the timbre space under evaluation are retrieved using the Euclidean distance between corresponding features vectors.
For this query, the precision is defined as:
$$
p@5_q = \frac{1}{5} \sum_{n=1}^5 \mathbf{1}_{D_q=D_n}
$$
The p@5 is the average $p@5_q$ computed over all queries.
\ja{This is perhaps better described in reverse, starting with $p@5_q$, then $p@5$, then $ap@5$?
Also, we define the precisions here as proportions (between zero and one), but later, we present them as percentages (between $0\%$ and $100\%$).}%

@@

The choice of the rank is chosen with the application in mind.
When searching for sounds that are similar to a query, the listener is likely to listen to up to $5$ proposals, rarely more.
For the same reason, more global metrics commonly considered for ranking tasks such as the mean average precision are not considered here.

@@

The processing pipeline is composed of two main building blocks.
The first block $B_1$ takes a digital audio recording as input and outputs one vector of features.
For this block, the main factors are the time support $B_1^t$ for computing the features and the type of features $B_1^f$.
We considered $B_1^t \in \{ 25, 128, 250, 500, 1000 \}$ ms and $B_1^f \in \{ \text{MFCCs}, \text{separable scattering}, \text{joint scattering} \}$.
The MFCCs are the standard mel-frequency cepstral coefficients with $40$ coefficients out of which we select the first $13$.
\ja{I'm not sure I like this notation.
First, why do we need to label the blocks ($B_1$ and $B_2$)?
We only have two of them.
Isn't it more straightforward to just call them ``the unsupervised part'' and ``the learned part'' or something like that?
We can also replace $B_1^t$ with something more straightforward, like $T$ (otherwise, it's a lot of bits to encode very little information).
Likewise, instead of $B_1^f$, we can just have $F$.}

@@

As found in many studies, log compression and median processing are beneficial post processing steps for the scattering.
\ja{References.}

For each recording, a set of features are computed for each time frame.
This set is averaged over time to give one feature vector per recording.
All features are standardized with zero mean and unit variance across the dataset.
\ja{Is this the ``per-feature Gaussianization'' we mention in the introduction?
Should give the reference in that case (and maybe explain a bit more in detail).}%

The second block $B_2$ considers one vector of features and the consensus clustering $D_c$ as inputs and another vector of features as output.
For this block, the main factor is the type of projection $B_2^p \in \{ \text{LDA}, \text{LMNN} \}$.
The LDA is the linear discriminant analysis \cite{} and the LMNN is the algorithm presented in Section \ref{sec:weighting}.
For this algorithm the neighborhood is set to $5$.
\ja{Again, I would prefer to ditch this $B_2$ notation.}

The complete experiments with all settings have been computed.
For the sake of clarity, only part of the results are discussed here.
\ja{What does this mean?}%
